---
title: "Ecudor Data Exploration"
author: "Sarah Sullivan"
output: html_document
---


```{r set_knitr_chunk_options, echo = F, warning=F, message=F}
rm(list = ls())

library(knitr)
library(markdown)
opts_chunk$set(message=FALSE, echo = F,
               warning = F,fig.height = 5, fig.width = 10, tidy = F)


#School
#opts_knit$set(root.dir = "H:/Ecuador/Code")
#Laptop
#opts_knit$set(root.dir = "/Users/sarahsullivan/Box Sync/Emory/Thesis/Code")
```

```{r libs}
library("ggplot2")
library("psych")
library("Hmisc")
library("reshape2")
library("plyr")
library("data.table")
library("geosphere")
```



```{r import}
#setwd("/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code")

#ecu<-read.table("/Users/sarahsullivan/Box Sync/NSF_Ecuador/Intervillage Datasets/Surv_red1.csv", sep = ",", quote = "", header=TRUE)
#save(ecu, file="EcuIndiv.RData")

#ecu17dis<-read.csv("/Users/sarahsullivan/Box Sync/Emory/Ecuador/DistanceData/17_SurvDataDist.csv")
#save(ecu17dis, file="/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code/Ecu17Dist.RData")

load("/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code/Ecu17Dist.RData")
load("/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code/EcuIndiv.RData")

```

```{r data_expl1}
describe(ecu)
describe(ecu$WEEK)
```

176 weeks
5897 unique individuals
21 Communities 
  Communities range in size from 28 to 872
  Number of people in each community varies between years, usually at three different timepoints (though that varies as well)
  Largest communities are: 
    number 5 (between 717 and 872 individuals), 
    followed by #4 with 445 to 553 individuals
  Smallest communities are:
    number 14 with 28-33 individuals
    followed by 1 with 52-54 and 12 with 30-78
  Potential Typical Communities
    13: median number of people (122), but the number of people varies highly, and given the transitory nature this seems like subsequent analyses will be difficult
    21: 121, 129, 152
    8: 136, 148, 156
    10: 107, 132, 135
  
1151 different households
  between 1 and 17 people in a household
Illness
  0.009202 or 0.92% sick (4944 observations)
    Fever 0.004612 or 0.46% had fever (2478 observations)
    Diarrhea 0.005079 or 0.51% (2729 observations)

```{r data_expl2}
ecu2<-unique(ecu[c("comunidad", "num_people_com")])
ecu3<-ecu2[order(ecu2$num_people_com),]
ecu4<-ecu[which(ecu$comunidad==8),]
ecu5<-ecu[order(ecu$WEEK),]
ecu5<-ecu5[!duplicated(ecu5$ID_individuo), ]
describe(ecu5)
```

Mean age at first data collection 21.64 (missing:604, over 10% problematic)
  Negative ages (-.9856)
    Need to figure out if there is data associated or if these are missing on other things 
Gender 52% male, 48% female (161 missing)
  1=male, 2=female

```{r data_expl3}
ecu6<-ecu[which(ecu$age<6),]
ecu7<-ecu6[!duplicated(ecu6$ID_individuo), ]
ecu8<-ecu[which(ecu$age<=5),]
ecu9<-ecu8[!duplicated(ecu8$ID_individuo), ]
```

Picking Villages:
  Check that there is variation within village chosen so that there is signal that we can assess
  Make sure there are enough children that its interesting
  Check on if its 5 and under or just under 5
  Check on the map
Crosstabs and Logistics to see what the drivers of disease are since our prevalences are low

To save computational time:
-save as RData set
-run on citrix
-drill down on subsets to first understand the data
  -single village
  -highest 10% of illness
  
Early questions:
-differential rates of dihr. across:
  -time, village age, sex, other relevant covariates
-id people frequently sick and see if the villge is sicker the next week 
-visualizing patterns in illness over time

Later Questions:
-village incidence to lag predict other villages
-distances for spatial analyses


```{r compare_villages}
ecu<- ecu[order(ecu$WEEK),] 
unique<-ecu[!duplicated(ecu$ID_individuo),]
vill8<-ecu[which(ecu$comunidad==8),]
uvil8<-vill8[!duplicated(vill8$ID_individuo),]
vill15<-ecu[which(ecu$comunidad==15),]
compop<-ecu[unique(ecu$comunidad),]

#compare
mean(ecu$DIARRHEA)
mean(unique$DIARRHEA)
mean(vill8$DIARRHEA)
mean(uvil8$DIARRHEA)

table(uvil8$WEEK)
#most people enrolled at first week, some enrollwd at later times
describe(vill8$ID_individuo)
#some weird high values, see if they mean anything

ecuid<-ecu$ID_individuo
ecuid<- ecuid[order(ecuid)] 
ecuidw<-ecuid[which(ecuid>8888000)]
ecuid8<-ecuid[which(ecuid>8888000 & ecuid<9999000)]
ecuid9<-ecuid[which(ecuid>=9999000)]
describe(ecuidw)
describe(ecuid8)
describe(ecuid9)

hist(ecu$WEEK, breaks=200)
hist(vill8$WEEK, breaks=200)

hist(vill15$WEEK, breaks=200)

par(mfrow=c(3,7))
wkhist<-by(ecu[,5], ecu$comunidad, function(x) hist(x,breaks=200))
sapply(wkhist, plot)


obs<-nrow(ecu)
indiv<-length(unique(ecu$ID_individuo))
obs/indiv

st1<-sum(compop$num_people_com)
obs/st1
#4614
```

2751 weird values with 8888 or 9999 before 3 digit code, 486 unique
*557 8888 lead, 119 unique
*2174 9999 lead, 367 unique
*Is there meaning in this?

Returning to looking at the week distribution

Visits per week
*Number of observations per week is usually around 3000 with the exception of the first week
*Number double for the first week (look into this),
*Seem to get most people most weeks
*sometimes half in the third year weeks 151-154 missing
*Weeks 45,46,50,51,151,152,153,154 are missing
*Looking at the distribution of samples from each week for each village, they are usually close to the max with some variations (90% by sight)
*The average number of visits per person x is 91.11 which is much smaller than the number of weeks collected (176). This appears to arise from 
*Still low and doesn't take into account new people

Need to reshape visit data wide (outcome=diarrhea)
  Can use reshape to find average number of visits per person

```{r reshape}
#Will work with vill8
Hmisc::describe(vill8)
v8wide <- dcast.data.table(data.table(vill8), ID_individuo + comunidad ~ WEEK, value.var=list("DIARRHEA"))
v8w<- as.data.frame(v8wide)
Hmisc::describe(v8w)

vis<-apply(v8wide[,c(3:166)], 1, function(x) sum(x, na.rm=T))
dat$comp<-ifelse(is.na(anymiss)==F, 1,0)


CD4w$numvis<-apply(CD4w, 1, function(x) (length(which(!is.na(x)))-2)/3)
CD4w$trt<-apply(CD4w[,17:23], 1, function(x) mean(x,na.rm = T))
table(CD4w$trt)
```


#Questions for Chris/Phil
-Is there any meaning in the strange identifiers?
--#2751 weird values with 8888 or 9999 before 3 digit code, 486 unique
--#557 8888 lead, 119 unique
--#2174 9999 lead, 367 unique

```{r village5}

ecu5a<-ecu[which(ecu$comunidad==5),]

ecu5<-ecu5a[order(ecu5a$WEEK),]
ecu5<-ecu5[!duplicated(ecu5$ID_individuo), ]
table(ecu5a$num_people_com)
describe(ecu5a)

#1125 individual people, maximum size at any one point is 872 (at least 253 people leave or are lost) 22%
```

```{r village4}

ecu4a<-ecu[which(ecu$comunidad==4),]

ecu4<-ecu4a[order(ecu4a$WEEK),]
ecu4<-ecu4[!duplicated(ecu4$ID_individuo), ]
table(ecu4a$num_people_com)
describe(ecu4a)
```
Village 4
*Max size: 553
*Unique: 734
*at least 181 missing, 25%


```{r village17}
ecu17a<-ecu[which(ecu$comunidad==17),]

ecu17a<-ecu17a[order(ecu17a$WEEK),]
ecu17<-ecu17a[!duplicated(ecu17a$ID_individuo), ]
table(ecu17a$num_people_com)
describe(ecu17a)

#Max: 471
#Unique: 557 
#at least 86 missing, but that's only 18%
```

Probably the best village since there is relatively little traffic in and out but there are still a good number of people 


```{r ecumove}

ecumov<-unique(ecu[c("comunidad", "ID_individuo")])
table(table(ecumov$ID_individuo, ecumov$comunidad))
```

Movement investigation
*there is a lot of movement but no people show up in two different villages: perhaps this is the weird 8888/9999 prefixes
*in a way its clean but we may be missing what is happening on an individual level
*very roughly over half of the time its only one person in a household who is sick at a time


```{r check8888dups}
load("/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code/EcuIndiv.RData")
ecua<-ecu[!duplicated(ecu$ID_individuo), ]
ecua<-ecua[order(ecua$ID_individuo),]
ecua$id2<-as.character(ecua$ID_individuo)
ecua$id3<-gsub("8888", "", ecua$id2, fixed=T)
ecua$id4<-gsub("9999", "", ecua$id3, fixed=T)

describe(ecua$id2)
describe(ecua$id3)
describe(ecua$id4)
ecua$id5<-as.numeric(ecua$id4)
describe(ecua$id5)
```

It appeared that the 8888s and 9999s overlapped, but they are in fact seperate individuals. Now checking to seee if this is because of a village change. It doesn't appear that we have any linked data for who these pople are since there's overlap, and there is no way for those IDs to be unique 


```{r invest8888}
ecu17<-ecu[which(ecu$comunidad==17),]

ecu17<-ecu17[order(ecu17$WEEK),]
ecu17a<-ecu17[!duplicated(ecu17$ID_individuo), ]
describe(ecu17a$ID_individuo)

ecu17a$id2<-as.character(ecu17a$ID_individuo)
ecu17a$id3<-gsub("8888", "", ecu17a$id2, fixed=T)
ecu17a$id4<-gsub("9999", "", ecu17a$id3, fixed=T)

describe(ecu17a$id2)
describe(ecu17a$id3)
describe(ecu17a$id4)
ecu17a$id5<-as.numeric(ecu17a$id4)
describe(ecu17a$id5)

table(ecua$id5,ecua$comunidad)

ecu77<-ecu[which(ecu$ID_individuo==9999077 | ecu$ID_individuo==77 | ecu$ID_individuo==8888077),]
```

These are all unique meaning that the 8888/9999 do not occur within their original village, ie 8888001 and 1 would not happen in the same village

77 is in both village 3 and 10, and since this should be unique instances it should have been a 9999/8888

Not the same people, since they're recorded in different villages (for about 5 weeks) at the same time. However 8888077 does have some missing info and is only present for a short time, so that might explain some things



```{r weekinvestigation}

table(table(ecu$ID_individuo))
hist(table(ecu$ID_individuo))

#can we cut those who are present less than 5 times?

ecua<-ecua[order(ecua$ID_individuo),]
wksperid<-ddply(ecu, .(ecu$ID_individuo),nrow)
ecua$wksperid<-wksperid$V1

hist(ecua$wksperid)
table(ecua$wksperid)
hist(ecua$WEEK[which(ecua$wksperid<=5)])

describe(ecua$wksperid[which(ecua$ID_individuo>=8888000)])
mean(ecua$DIARRHEA[which(ecua$ID_individuo>=8888000)])
mean(ecu$DIARRHEA[which(ecu$ID_individuo>=8888000)])
hist()

describe(ecua$wksperid[which(ecua$ID_individuo<8888000)])
mean(ecua$DIARRHEA[which(ecua$ID_individuo<8888000)])
mean(ecu$DIARRHEA[which(ecu$ID_individuo<8888000)])
mean(ecu$DIARRHEA)
1-nrow(ecu[which(ecu$ID_individuo<8888000),])/nrow(ecu)
1-mean(ecu$DIARRHEA[which(ecu$ID_individuo<8888000)])/mean(ecu$DIARRHEA)

```


```{r village17}
ecu17<-ecu[which(ecu$comunidad==17),]
ecu17a<-ecua[which(ecua$comunidad==17),]


ecuad<-ecu[!duplicated(ecu$ID_casa), ]
ecu17ad<-ecu17[!duplicated(ecu17$ID_casa), ]
ecu17ad<-ecu17ad[order(ecu17ad$ID_casa),]

```

```{r distmatChris}
ecu17dh<-ecu17dis[!duplicated(ecu17dis$ID_casa), ]

#names(ecu17dh)

distm<-ecu17dh[,c(2,47:195)]

matnames<-names(distm)
matnames<-sort(matnames)
matnames[151]<-"DISTANCE...1"
head(matnames)

#Could pull out stuff if I need to but I will save that for another time since other methods willbe faster


```

```{r distmatSarah}

earth.dist <- function (long1, lat1, long2, lat2)
{
rad <- pi/180
a1 <- lat1 * rad
a2 <- long1 * rad
b1 <- lat2 * rad
b2 <- long2 * rad
dlon <- b2 - a2
dlat <- b1 - a1
a <- (sin(dlat/2))^2 + cos(a1) * cos(b1) * (sin(dlon/2))^2
c <- 2 * atan2(sqrt(a), sqrt(1 - a))
R <- 6378.145
d <- R * c
return(d)
}
#https://conservationecology.wordpress.com/2013/06/30/distance-between-two-points-in-r/

#GIS hh 1 and 2
1000*earth.dist(-79.06954, 0.7993030, -79.06957,0.7992739)
# 4.652355
###Not Accurate Don't Use


#from Chris
distm[1,134]
# 4.584152

#Utilizes an elipsoidal method to calculate distance 
gdistm<-distm(ecu17ad[,c("hh_lon","hh_lat")], fun=distVincentyEllipsoid)

distcw<-data.frame(ID_casa=ecu17ad$ID_casa,mat_ID=c(1:107),ID_casa2=ecu17dh$ID_casa, GIS_FID=ecu17dh$GIS_FID) 

#make sure that the ID_casa matches up for both datasets
distcw$ID_casa==distcw$ID_casa2
  #succes
  
#write.csv(gdistm, file="/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code/DistanceMatrix17.csv")
#write.csv(distcw, file="/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code/DistCrosswalk17.csv")


```

```{r aggregateWkCases}
weekcases17<-aggregate(ecu17[,"DIARRHEA"], by=list(ecu17$WEEK), FUN=sum)
wkcases17<-rbind(weekcases17[1:113,],c(114,NA), weekcases17[114:116,], c(118,NA), weekcases17[117:120,], c(123,NA),weekcases17[121:174,])
wkcs17<-wkcases17$x
```


```{r aggWkCaseHouse}

aggwks17<-aggregate(ecu17[,"DIARRHEA"], by=list(ecu17$WEEK, ecu17$ID_casa), FUN=sum)

hhweekcase17<-dcast(aggwks17, Group.2~Group.1, value.var="x")
hhwkcase17<-cbind(hhweekcase17[,2:114],rep(NA, times=107), hhweekcase17[,115:117], rep(NA, times=107), hhweekcase17[,118:121], rep(NA, times=107),hhweekcase17[,122:175])

#write.csv(hhwkcase17, file="/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code/WeeklyHouseholdCases17.csv")

```

###Matrices (cleaned and identified by index for week and household):

Dataset | Description | Dimmensions
------------- | ------------- | -------------
wkcs7 | total cases per week | 1x177
gdistm |      distance matrix for households | 107x107
hhwkcase17 |  number of cases within households by week | 107x177
distcw |      crosswalk betweek household ids and household indexes | 4x107

