---
title: "Ecudor Data Exploration"
author: "Sarah Sullivan"
output: html_document
---


```{r set_knitr_chunk_options, echo = F, warning=F, message=F}
rm(list = ls())

library(knitr)
library(markdown)
opts_chunk$set(message=FALSE, echo = F,
               warning = F,fig.height = 5, fig.width = 10, tidy = F)


#School
#opts_knit$set(root.dir = "H:/Ecuador/Code")
#Laptop
#opts_knit$set(root.dir = "/Users/sarahsullivan/Box Sync/Emory/Thesis/Code")
```

```{r libs}
library("ggplot2")
library("psych")
library("Hmisc")
library("reshape2")
library("plyr")
library("data.table")
```



```{r import}
setwd("/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code")

ecu<-read.table("/Users/sarahsullivan/Box Sync/NSF_Ecuador/Intervillage Datasets/Surv_red1.csv", sep = ",", quote = "", header=TRUE)
save(ecu, file="EcuIndiv.RData")

load("/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code/EcuIndiv.RData")

```

```{r data_expl}
describe(ecu)
describe(ecu$WEEK)
#176 weeks
#5897 unique individuals
#21 Communities 
  #Communities range in size from 28 to 872
  #Number of people in each community varies between years, usually at three different timepoints (though that varies as well)
  #Largest communities are: 
    #number 5 (between 717 and 872 individuals), 
    #followed by #4 with 445 to 553 individuals
  #Smallest communities are:
    #number 14 with 28-33 individuals
    #followed by 1 with 52-54 and 12 with 30-78
  #Potential Typical Communities
    #13: median number of people (122), but the number of people varies highly, and given the transitory nature this seems like subsequent analyses will be difficult
    #21: 121, 129, 152
    #8: 136, 148, 156
    #10: 107, 132, 135
  
#1151 different households
  #between 1 and 17 people in a household
#Illness
  #0.009202 or 0.92% sick (4944 observations)
    #Fever 0.004612 or 0.46% had fever (2478 observations)
    #Diarrhea 0.005079 or 0.51% (2729 observations)


ecu2<-unique(ecu[c("comunidad", "num_people_com")])
ecu3<-ecu2[order(ecu2$num_people_com),]
ecu4<-ecu[which(ecu$comunidad==8),]
ecu5<-ecu[order(ecu$WEEK),]
ecu5<-ecu5[!duplicated(ecu5$ID_individuo), ]
describe(ecu5)

#Mean age at first data collection 21.64 (missing:604, over 10% problematic)
  #Negative ages (-.9856)
    #Need to figure out if there is data associated or if these are missing on other things 
#Gender 52% male, 48% female (161 missing)
  #1=male, 2=female

ecu6<-ecu[which(ecu$age<6),]
ecu7<-ecu6[!duplicated(ecu6$ID_individuo), ]
ecu8<-ecu[which(ecu$age<=5),]
ecu9<-ecu8[!duplicated(ecu8$ID_individuo), ]

#Picking Villages:
  #Check that there is variation within village chosen so that there is signal that we can assess
  #Make sure there are enough children that its interesting
  #Check on if its 5 and under or just under 5
  #Check on the map
#Crosstabs and Logistics to see what the drivers of disease are since our prevalences are low
```


To save computational time:
-save as RData set
-run on citrix
-drill down on subsets to first understand the data
  -single village
  -highest 10% of illness
  
Early questions:
-differential rates of dihr. across:
  -time, village age, sex, other relevant covariates
-id people frequently sick and see if the villge is sicker the next week 
-visualizing patterns in illness over time

Later Questions:
-village incidence to lag predict other villages
-distances for spatial analyses


```{r compareViltoAll}
ecu<- ecu[order(ecu$WEEK),] 
unique<-ecu[!duplicated(ecu$ID_individuo),]
vill8<-ecu[which(ecu$comunidad==8),]
uvil8<-vill8[!duplicated(vill8$ID_individuo),]
vill15<-ecu[which(ecu$comunidad==15),]
compop<-ecu[unique(ecu$comunidad),]

#compare
mean(ecu$DIARRHEA)
mean(unique$DIARRHEA)
mean(vill8$DIARRHEA)
mean(uvil8$DIARRHEA)

table(uvil8$WEEK)
#most people enrolled at first week, some enrollwd at later times
describe(vill8$ID_individuo)
#some weird high values, see if they mean anything

ecuid<-ecu$ID_individuo
ecuid<- ecuid[order(ecuid)] 
ecuidw<-ecuid[which(ecuid>8888000)]
ecuid8<-ecuid[which(ecuid>8888000 & ecuid<9999000)]
ecuid9<-ecuid[which(ecuid>=9999000)]
describe(ecuidw)
describe(ecuid8)
describe(ecuid9)

#2751 weird values with 8888 or 9999 before 3 digit code, 486 unique
#557 8888 lead, 119 unique
#2174 9999 lead, 367 unique

#Is there meaning in this?

#returning to looking at the week distribution

hist(ecu$WEEK, breaks=200)
#Number of observations per week is usually around 3000 with the exception of the first week

hist(vill8$WEEK, breaks=200)
#Number double for the first week (look into this),
#Seem to get most people most weeks, sometimes half in the third year weeks 151-154 missing
#Weeks 45,46,50,51,151,152,153,154 are missing

hist(vill15$WEEK, breaks=200)

par(mfrow=c(3,7))
wkhist<-by(ecu[,5], ecu$comunidad, function(x) hist(x,breaks=200))
sapply(wkhist, plot)

sapply()

#Looking at the distribution of samples from each week for each village, they are usually close to the max with some variations (90% by sight)
obs<-nrow(ecu)
indiv<-length(unique(ecu$ID_individuo))
obs/indiv
#The average number of visits per person x is 91.11 which is much smaller than the number of weeks collected (176). This appears to arise from 


st1<-sum(compop$num_people_com)
obs/st1
#still low and doesn't take into account new people
#4614



```

#Need to reshape visit data wide (outcome=diarrhea)
###Can use reshape to find average number of visits per person

```{r reshape}
#Will work with vill8
Hmisc::describe(vill8)
v8wide <- dcast.data.table(data.table(vill8), ID_individuo + comunidad ~ WEEK, value.var=list("DIARRHEA"))
v8w<- as.data.frame(v8wide)
Hmisc::describe(v8w)

vis<-apply(v8wide[,c(3:166)], 1, function(x) sum(x, na.rm=T))
dat$comp<-ifelse(is.na(anymiss)==F, 1,0)


CD4w$numvis<-apply(CD4w, 1, function(x) (length(which(!is.na(x)))-2)/3)
CD4w$trt<-apply(CD4w[,17:23], 1, function(x) mean(x,na.rm = T))
table(CD4w$trt)
```


#Questions for Chris/Phil
-Is there any meaning in the strange identifiers?
--#2751 weird values with 8888 or 9999 before 3 digit code, 486 unique
--#557 8888 lead, 119 unique
--#2174 9999 lead, 367 unique

```{r v5}

ecu5a<-ecu[which(ecu$comunidad==5),]

ecu5<-ecu5a[order(ecu5a$WEEK),]
ecu5<-ecu5[!duplicated(ecu5$ID_individuo), ]
table(ecu5a$num_people_com)
describe(ecu5a)

#1125 individual people, maximum size at any one point is 872 (at least 253 people leave or are lost) 22%
```

```{r v5}

ecu4a<-ecu[which(ecu$comunidad==4),]

ecu4<-ecu4a[order(ecu4a$WEEK),]
ecu4<-ecu4[!duplicated(ecu4$ID_individuo), ]
table(ecu4a$num_people_com)
describe(ecu4a)

#Max size: 553
#Unique: 734
#at least 181 missing, 25%
```

```{r v17}
ecu17a<-ecu[which(ecu$comunidad==17),]

ecu17a<-ecu17a[order(ecu17a$WEEK),]
ecu17<-ecu17[!duplicated(ecu17$ID_individuo), ]
table(ecu17a$num_people_com)
describe(ecu17a)

Max: 471
Unique: 557 
#at least 86 missing, but that's only 18%
```

```{r ecumove}

ecumov<-unique(ecu[c("comunidad", "ID_individuo")])
table(table(ecumov$ID_individuo, ecumov$comunidad))


#there is a lot of movement but no people show up in two different villages: perhaps this is the weird 8888/9999 prefixes

#in a way its clean but we may be missing what is happening on an individual level

# very roughly 70% of the time its only one person in a household who is sick at a time

```

```{r}
load("/Users/sarahsullivan/Box Sync/Emory/Ecuador/Code/EcuIndiv.RData")
ecua<-ecu[!duplicated(ecu$ID_individuo), ]
ecua<-ecua[order(ecua$ID_individuo),]
ecua$id2<-as.character(ecua$ID_individuo)
ecua$id3<-gsub("8888", "", ecua$id2, fixed=T)
ecua$id4<-gsub("9999", "", ecua$id3, fixed=T)

describe(ecua$id2)
describe(ecua$id3)
describe(ecua$id4)
ecua$id5<-as.numeric(ecua$id4)
describe(ecua$id5)

#The 8888 and 9999 leading IDs are indeed already in the dataset with their base IDs. Now checking to seee if this is because of a village change

```

```{r}
ecu17<-ecu[which(ecu$comunidad==17),]

ecu17<-ecu17[order(ecu17$WEEK),]
ecu17a<-ecu17[!duplicated(ecu17$ID_individuo), ]
describe(ecu17a$ID_individuo)

ecu17a$id2<-as.character(ecu17a$ID_individuo)
ecu17a$id3<-gsub("8888", "", ecu17a$id2, fixed=T)
ecu17a$id4<-gsub("9999", "", ecu17a$id3, fixed=T)

describe(ecu17a$id2)
describe(ecu17a$id3)
describe(ecu17a$id4)
ecu17a$id5<-as.numeric(ecu17a$id4)
describe(ecu17a$id5)

#These are all unique meaning that the 8888/9999 do not occur within their original village

table(ecua$id5,ecua$comunidad)
  #77 is in both village 3 and 10, and since this should be unique instances it should have been a 9999/8888


ecu77<-ecu[which(ecu$ID_individuo==9999077 | ecu$ID_individuo==77 | ecu$ID_individuo==8888077),]
#Not the same people, since they're recorded in different villages (for about 5 weeks) at the same time. However 8888077 does have some missing info and is only present for a short time, so that might explain some things

```


```{r weekinvestigation}

table(table(ecu$ID_individuo))
hist(table(ecu$ID_individuo))

#can we cut those who are present less than 5 times

ecua<-ecua(sort())
wksperid<-ddply(ecu, .(ecu$ID_individuo),nrow)
ecua$wksperid<-wksperid$V1

hist(ecua$wksperid)
table(ecua$wksperid)
hist(ecua$WEEK[which(ecua$wksperid<=5)])



describe(ecua$wksperid[which(ecua$ID_individuo>=8888000)])
mean(ecua$DIARRHEA[which(ecua$ID_individuo>=8888000)])
mean(ecu$DIARRHEA[which(ecu$ID_individuo>=8888000)])
hist()

describe(ecua$wksperid[which(ecua$ID_individuo<8888000)])
mean(ecua$DIARRHEA[which(ecua$ID_individuo<8888000)])
mean(ecu$DIARRHEA[which(ecu$ID_individuo<8888000)])
mean(ecu$DIARRHEA)
1-nrow(ecu[which(ecu$ID_individuo<8888000),])/nrow(ecu)
1-mean(ecu$DIARRHEA[which(ecu$ID_individuo<8888000)])/mean(ecu$DIARRHEA)



```


```{r households}





```
